Canonical rings of rational divisors on projective space and Hirzebruch surfaces

### 2015-07-11 surfaces-write-up/surface-write-up.tex
surfaces title change, curves abs one-word edit
add def lower approx, first half 1pt pf, add pic

\title{Spin canonical rings of log stacky curves}

@@

@@

\begin{thm}
\label{thm:proj-one-point}
Let $X = \bp^m$ with $\{x_0, \ldots, x_m\}$ a coordinate system for
$\bp^m$. Let $D = \frac{a}{b} H$, with $a, b \in \bz$ and $H =
V(x_k)$ a hyperplane of $\bp^m$, be a divisor on $\bp^m$. Let
\[
	\left\lfloor \frac{a}{b} \right\rfloor = \frac{c_0}{d_0} < \frac{c_1}{d_1} <
	\ldots < \frac{c_r}{d_r} = \frac{a}{b}
\]

\noindent
be the best lower approximations of $\frac{a}{b}$. Then the
canonical ring
\[
	R_D := \bigoplus_{d \geq 0} H^0(\bp^m, \lfloor dD \rfloor)
\]

\noindent
has a minimal presentation consisting of the $\sum_{i = 0}^{r}
{{m + i} \choose {i}}$ generators $f_{\vec{v}} := \frac{u^{d_i}
x^{\vec{v}}}{x_k^{c_i}}$ where $\vec{v} \in \bz_{\geq 0}^{m + 1}$
such that $\sum_{j = 0}^{m} v_j = c_i$ and $z$ \todo{how many?}
relations of the form
\[
	x^{\vec{u}} x^{\vec{v}} = x^{\vec{w}} x^{\vec{z}}
\]

@@

\begin{proof}
We know $f_0$ is a minimal generator in degree $1$. If $x^{\vec{v}}$
were not minimal for $\deg \vec{v} = c_i > 0$ for some $i > 0$, then
$f_{\vec{v}} = f_{\vec{w}} f_{\vec{z}}$ for some $f_{\vec{w}},
f_{\vec{z}} \in R_D$ with $\deg \vec{w}, \deg \vec{z} < \deg \vec{v}$.
\end{proof}

@@

\section{Canonical Rings on \texorpdfstring{$F_k$}{}}
\begin{defn}
Suppose $D = \sum_{i=1}^m \alpha_i C_i \in \mathbb{Q} \otimes \di(F_k)$.  Then define 
\begin{equation}\label{eqn:define-T=(D)}
	T_=(D) = \left\{i \in \{1, \ldots, m\}: d_i \sum_{k=1}^m d_k' \alpha_k = d_i' \sum_{k=1}^m d_k \alpha_k \right\},
\end{equation}
\begin{equation}\label{eqn:define-T+(D)}
	T_+(D) = \left\{ i \in \{1, \ldots, m\}:  \vphantom{\sum_{k=1}^m} d_i \sum_{k=1}^m \alpha_k d_k' > d_i' \sum_{k=1}^m \alpha_k d_k \right\},
\end{equation}
and
\begin{equation}\label{eqn:define-T-(D)}
	T_-(D) = \left\{ j \in \{1, \ldots, m\}: d_j \sum_{k=1}^m \alpha_k d_k' < d_j' \sum_{k=1}^m \alpha_k d_k s \right\}.
\end{equation}
\end{defn}

>>>>

\title{Canonical rings of rational divisors on projective space and Hirzebruch surfaces}

@@

\begin{defn}
\label{defn:lower-approximation}
If $\frac{\alpha}{\beta}\in \bq$, then $\frac{c}{k}$ (written in
reduced form with $c \in \bn, k \in \bn$) is a {\bf{best lower
approximation}} to $\frac{\alpha}{\beta}$ if there does not exist
any $\frac{c'}{k'}\in \bq$ with $0 < k'<k$ and $\frac{c}{k} \le
\frac{c'}{k'}$.
\end{defn}

Moreover, if $\frac{\alpha}{\beta}\ge 0$, then the non-negative
best lower approximations of $\frac{\alpha}{\beta}$ form a finite
sequence
\[
	0 = \frac{c_0}{k_0} < \frac{c_1}{k_1} < \ldots < \frac{c_r}{k_r} = \frac{\alpha}{\beta}.
\]
Figure \ref{fig:s14/5-lattice} gives a pictorial representation of
the positive best lower approximations of $\frac{\alpha}{\beta} =
\frac{14}{5}$.

\begin{figure}
\includegraphics{pics/spin-lower-approximations-pic-pics.pdf}
\caption{This figure shows each non-negative best lower
approximation of $\frac{14}{5}.$ Each ``$\bullet$'' denotes a best
lower approximation and each ``$\circ$'' denotes a lattice point
below $5y=14x$ which is not a best lower approximation.  Note that
the non-negative best lower approximations generate the monoid of
lattice points in the first quadrant satisfying  $5y \le 14x$, with
the operation $(a_1, b_1)(a_2, b_2)\mapsto (a_1 + a_2, b_1 + b_2)$.}
\label{fig:s14/5-lattice}
\end{figure}

@@

\begin{thm}
\label{thm:proj-one-point}
Let $X = \bp^m$ with $\{x_0, \ldots, x_m\}$ a coordinate system for
$\bp^m$. Let $D = \alpha H$, with $\alpha = \frac{a}{b} \in \bq$
and $H = V(x_k)$ a hyperplane of $\bp^m$, be a divisor on $\bp^m$.
Let
\[
	\left\lfloor \frac{a}{b} \right\rfloor = \frac{c_0}{d_0} <
	\frac{c_1}{d_1} < \ldots < \frac{c_r}{d_r} = \frac{a}{b}
\]

\noindent
be the best lower approximations of $\alpha$. Then the
canonical ring
\[
	R_D := \bigoplus_{d \geq 0} H^0(\bp^m, \lfloor dD \rfloor)
\]

\noindent
has a minimal presentation consisting of the $\sum_{i = 0}^{r}
{{m + i} \choose {i}}$ many generators $f_{d_i, \vec{v}} :=
\frac{u^{d_i} x^{\vec{v}}} {x_k^{c_i}}$ where $\vec{v} \in
\bz_{\geq 0}^{m + 1}$ such that $\sum_{j = 0}^{m} v_j = c_i$ and $z$ \todo{how many?}
many relations of the form
\[
	x^{\vec{u}} x^{\vec{v}} = x^{\vec{w}} x^{\vec{z}}
\]

@@

\begin{proof}
We know $\{f_{0, \vec{u}} : \deg \vec{u} = \left\lfloor \frac{a}{b}
\right\rfloor \}$ are minimal generators in degree $1$. If
$f_{d_i, \vec{v}}$ were not minimal for $\deg \vec{v} = c_i$ for
some $i > 0$, then $f_{d_i, \vec{v}} = f_{s, \vec{w}}
f_{t, \vec{z}}$ for some $s, t \in \bn$ and $\vec{w}, \vec{z} \in
\bz_{\geq 0}^{m + 1}$. In particular, we see that 
\begin{align*}
	&d_i = s + t \\
	&c_i = \deg \vec{v} = \deg \vec{w} + \deg \vec{z}
\end{align*}

\noindent
so $\frac{c_i}{d_i}$ is the mediant of the rational numbers
$\frac{\deg \vec{w}}{s}, \frac{\deg \vec{z}}{t}$ which both
do not exceed $\alpha$. One of $\frac{\deg \vec{w}}{s},
\frac{\deg \vec{z}}{t}$ must be at least $\frac{c_i}{d_i}$
and $s$ and $t$ are both less than $d_i$, so we have a contradiction 
of $\frac{c_i}{d_i}$ being a best lower approximation of $\alpha$
(see Definition ~\ref{defn:lower-approximation}).

Now we show that no further generators are necessary. Let $\vec{x}_i
:= (d_i, c_i) \in \bz^2$ a lattice point.

@@

\section{Canonical Rings on Hirzebruch Surfaces}
\begin{defn}
Suppose $D = \sum_{i=1}^m \alpha_i C_i \in \mathbb{Q} \otimes \di(F_k)$.  Then define 
\begin{equation}\label{eqn:define-T=(D)}
	T_=(D) = \left\{i \in \{1, \ldots, m\}: d_i \sum_{k=1}^m d_k' \alpha_k = d_i' \sum_{k=1}^m d_k \alpha_k \right\},
\end{equation}
\begin{equation}\label{eqn:define-T+(D)}
	T_+(D) = \left\{ i \in \{1, \ldots, m\}:  \vphantom{\sum_{k=1}^m} d_i \sum_{k=1}^m \alpha_k d_k' > d_i' \sum_{k=1}^m \alpha_k d_k \right\},
\end{equation}
and
\begin{equation}\label{eqn:define-T-(D)}
	T_-(D) = \left\{ j \in \{1, \ldots, m\}: d_j \sum_{k=1}^m \alpha_k d_k' < d_j' \sum_{k=1}^m \alpha_k d_k s \right\}.
\end{equation}
\end{defn}
